---
title: "Reconcile PAC reviews"
output: html_notebook
---

# Load data and packages

```{r}
library(dplyr)
library(readr)
library(stringr)
```


```{r}
pac_review <- read_csv("04_inputs/pac_manual_review.csv",
                                 col_types = cols(.default = "c", under_5k_amount = "d")) # manual review feedback 

# add paths to any additional PAC review files here (oldest -> newest)
# each will be merged into pac_review in order, with conflicts flagged.
supplemental_review_paths <- c(
  "04_inputs/pac_manual_review_09-08-26.csv",
  "04_inputs/pac_review_queue_w_review_09-25-26.csv"
)

supplemental_reviews <- lapply(supplemental_review_paths, function(p) {
  read_csv(p, col_types = cols(.default = "c", under_5k_amount = "d")) %>%
    mutate(review_source = basename(p), pac_id = as.character(pac_id))
}) %>% bind_rows()
```

```{r}
supplemental_review_paths
```



### combine pac review into one dataframe

Major donors are flagged via `industry_override_under_5k = "Major Donors/Megadonors"` OR "major donor" in notes; they remain in pac_review with `is_major_donor = TRUE` so 0103 can suppress them from future review queues.

```{r}
# union subcategory labels that map to the broader "Unions" override in newer reviews.
union_subcategory_overrides <- c(
  "Misc Unions", "Unions", "Public Safety Unions", "Public Sector Unions",
  "Social Services Unions", "Trade Unions"
)

pac_review <- pac_review %>% mutate(pac_id = as.character(pac_id))

all_major_donors     <- list()
all_new_pending      <- list()
all_id_conflicts     <- list()
all_name_conflicts   <- list()
all_pending_resolved <- list()

for (src in unique(supplemental_reviews$review_source)) {
  supp <- supplemental_reviews %>% filter(review_source == src)
  cat(sprintf("\n=== Merging: %s ===\n", src))

  # 1. flag major donors (by notes OR by override)
  major_donor_rows <- supp %>%
    filter(
      grepl("major donor", notes, ignore.case = TRUE) |
      trimws(coalesce(industry_override_under_5k, "")) == "Major Donors/Megadonors"
    )
  all_major_donors[[src]] <- major_donor_rows %>%
    mutate(review_source = src) %>%
    select(review_source, pac_id, Contributor.Name, industry_override_under_5k, notes)
  # notes-based major donors: standardize override to "Major Donors/Megadonors" so they're
  # consistently identifiable; include all rows in supp_clean so they enter pac_review
  supp_clean <- supp %>%
    mutate(industry_override_under_5k = if_else(
      grepl("major donor", notes, ignore.case = TRUE),
      "Major Donors/Megadonors",
      industry_override_under_5k
    ))

  # 2. pendings pacs
  new_pending <- supp_clean %>%
    filter(industry_override_under_5k == "PENDING" |
           unresolvable_reason == "additional_review_req")
  all_new_pending[[src]] <- new_pending %>%
    mutate(review_source = src) %>%
    select(review_source, pac_id, Contributor.Name,
           industry_override_under_5k, unresolvable_reason)

  # 3. fill pac_id for no-id rows where reviewer coded one (notes = "coded id" or "coded id possible")
  coded_id_rows <- supp_clean %>%
    filter(grepl("coded id", notes, ignore.case = TRUE),
           !is.na(pac_id), pac_id != "") %>%
    distinct(Contributor.Name, .keep_all = TRUE) %>%
    select(Contributor.Name, new_pac_id = pac_id)

  if (nrow(coded_id_rows) > 0) {
    pac_review <- pac_review %>%
      left_join(coded_id_rows, by = "Contributor.Name") %>%
      mutate(pac_id = if_else(
        (is.na(pac_id) | pac_id == "") & !is.na(new_pac_id),
        new_pac_id, pac_id
      )) %>%
      select(-new_pac_id)
    cat(sprintf("  Filled pac_id for %d previously no-id row(s).\n", nrow(coded_id_rows)))
  }

  # 4. detect conflicts (flagged for awareness; new review wins on all overlaps)
  # deduplicate to one override per pac_id/name before joining to avoid many-to-many warnings
  old_id_overrides <- pac_review %>%
    filter(!is.na(pac_id), pac_id != "",
           !is.na(industry_override_under_5k), industry_override_under_5k != "") %>%
    distinct(pac_id, override_old = industry_override_under_5k)

  new_id_overrides <- supp_clean %>%
    filter(!is.na(pac_id), pac_id != "",
           !is.na(industry_override_under_5k), industry_override_under_5k != "") %>%
    distinct(pac_id, override_new = industry_override_under_5k)

  id_conflicts <- old_id_overrides %>%
    inner_join(new_id_overrides, by = "pac_id", relationship = "many-to-many") %>%
    filter(
      override_old != override_new,
      override_old != "PENDING",
      !(override_old %in% union_subcategory_overrides & override_new == "Unions")
    ) %>%
    distinct() %>%
    mutate(review_source = src)
  all_id_conflicts[[src]] <- id_conflicts

  old_name_overrides <- pac_review %>%
    filter(is.na(pac_id) | pac_id == "",
           !is.na(industry_override_under_5k), industry_override_under_5k != "") %>%
    distinct(Contributor.Name, override_old = industry_override_under_5k)

  new_name_overrides <- supp_clean %>%
    filter(is.na(pac_id) | pac_id == "",
           !is.na(industry_override_under_5k), industry_override_under_5k != "") %>%
    distinct(Contributor.Name, override_new = industry_override_under_5k)

  name_conflicts <- old_name_overrides %>%
    inner_join(new_name_overrides, by = "Contributor.Name", relationship = "many-to-many") %>%
    filter(
      override_old != override_new,
      override_old != "PENDING",
      !(override_old %in% union_subcategory_overrides & override_new == "Unions")
    ) %>%
    distinct() %>%
    mutate(review_source = src)
  all_name_conflicts[[src]] <- name_conflicts

  cat(sprintf("  Major donors: %d | PENDING/add'l review: %d | pac_id conflicts: %d | name conflicts: %d\n",
              nrow(major_donor_rows), nrow(new_pending),
              nrow(id_conflicts), nrow(name_conflicts)))

  # 5. Apply new review's overrides (new review wins for all overlapping rows)
  # - industry_override_under_5k: updated where new review has a valid (non-empty, non-PENDING) value
  # - fec_id / is_federal_pac: updated for ALL overlapping rows (new review is more reliable)
  # - unresolvable_reason: cleared for rows that now have a valid override, so they are
  #   not incorrectly skipped by skip_pac_ids downstream

  pending_pac_ids_before <- pac_review %>%
    filter(!is.na(pac_id), pac_id != "", industry_override_under_5k == "PENDING") %>%
    pull(pac_id) %>% unique()

  # id-based: industry override — only where new review has a valid value
  new_id_override <- supp_clean %>%
    filter(!is.na(pac_id), pac_id != "",
           !is.na(industry_override_under_5k),
           industry_override_under_5k != "",
           industry_override_under_5k != "PENDING") %>%
    group_by(pac_id) %>% slice(1) %>% ungroup() %>%
    select(pac_id, new_override = industry_override_under_5k)

  # id-based: federal fields — all overlapping rows regardless of override status
  new_id_federal_base <- supp_clean %>%
    filter(!is.na(pac_id), pac_id != "") %>%
    group_by(pac_id) %>% slice(1) %>% ungroup()
  fed_cols <- intersect(c("fec_id", "is_federal_pac"), names(new_id_federal_base))
  new_id_federal <- new_id_federal_base %>%
    select(pac_id, all_of(fed_cols)) %>%
    rename_with(~ paste0("new_", .), all_of(fed_cols))

  pac_review <- pac_review %>%
    left_join(new_id_override, by = "pac_id") %>%
    mutate(industry_override_under_5k = if_else(
      !is.na(pac_id) & pac_id != "" & !is.na(new_override),
      new_override, industry_override_under_5k
    )) %>%
    select(-new_override) %>%
    left_join(new_id_federal, by = "pac_id")

  if ("new_fec_id" %in% names(pac_review))
    pac_review <- pac_review %>%
      mutate(fec_id = if_else(!is.na(pac_id) & pac_id != "" & !is.na(new_fec_id),
                              new_fec_id, fec_id)) %>%
      select(-new_fec_id)
  if ("new_is_federal_pac" %in% names(pac_review))
    pac_review <- pac_review %>%
      mutate(is_federal_pac = if_else(!is.na(pac_id) & pac_id != "" & !is.na(new_is_federal_pac),
                                      new_is_federal_pac, is_federal_pac)) %>%
      select(-new_is_federal_pac)

  # name-based: industry override only (federal fields not reliable for name-matched rows)
  new_name_override <- supp_clean %>%
    filter(is.na(pac_id) | pac_id == "",
           !is.na(industry_override_under_5k),
           industry_override_under_5k != "",
           industry_override_under_5k != "PENDING") %>%
    group_by(Contributor.Name) %>% slice(1) %>% ungroup() %>%
    select(Contributor.Name, new_override = industry_override_under_5k)

  pac_review <- pac_review %>%
    left_join(new_name_override, by = "Contributor.Name") %>%
    mutate(industry_override_under_5k = if_else(
      (is.na(pac_id) | pac_id == "") & !is.na(new_override),
      new_override, industry_override_under_5k
    )) %>%
    select(-new_override)

  # clear unresolvable_reason = "additional_review_req" for rows that now have a valid override
  pac_review <- pac_review %>%
    mutate(unresolvable_reason = if_else(
      !is.na(industry_override_under_5k) &
      industry_override_under_5k != "" &
      industry_override_under_5k != "PENDING" &
      coalesce(unresolvable_reason, "") == "additional_review_req",
      NA_character_,
      unresolvable_reason
    ))

  # track resolved pending pacs
  all_pending_resolved[[src]] <- pac_review %>%
    filter(pac_id %in% pending_pac_ids_before,
           !is.na(industry_override_under_5k),
           industry_override_under_5k != "PENDING",
           industry_override_under_5k != "") %>%
    distinct(pac_id, override_new = industry_override_under_5k) %>%
    mutate(review_source = src)

  cat(sprintf("  Applied new overrides to overlapping rows (%d PENDING resolved).\n",
              nrow(all_pending_resolved[[src]])))

  # 6. append new rows
  old_pac_ids_now <- pac_review %>% filter(!is.na(pac_id), pac_id != "") %>% pull(pac_id) %>% unique()
  old_names_noid  <- pac_review %>% filter(is.na(pac_id) | pac_id == "") %>% pull(Contributor.Name) %>% unique()

  # for no-id rows, also skip if any name_variant already appears in old review
  new_rows_noid <- supp_clean %>% filter(is.na(pac_id) | pac_id == "")
  if ("name_variants" %in% names(new_rows_noid) && nrow(new_rows_noid) > 0) {
    new_rows_noid <- new_rows_noid %>%
      rowwise() %>%
      mutate(covered_by_variant = !is.na(name_variants) && name_variants != "" &&
               any(trimws(strsplit(name_variants, "\\|")[[1]]) %in% old_names_noid)) %>%
      ungroup() %>%
      filter(!Contributor.Name %in% old_names_noid, !covered_by_variant) %>%
      select(-covered_by_variant)
  } else {
    new_rows_noid <- new_rows_noid %>% filter(!Contributor.Name %in% old_names_noid)
  }

  new_rows <- bind_rows(
    supp_clean %>% filter(!is.na(pac_id), pac_id != "", !pac_id %in% old_pac_ids_now),
    new_rows_noid
  )

  for (col in intersect(names(new_rows), names(pac_review))) {
    if (is.numeric(pac_review[[col]]) && !is.numeric(new_rows[[col]]))
      new_rows[[col]] <- as.numeric(new_rows[[col]])
    else if (is.character(pac_review[[col]]) && !is.character(new_rows[[col]]))
      new_rows[[col]] <- as.character(new_rows[[col]])
  }

  pac_review <- bind_rows(pac_review, new_rows)
  cat(sprintf("  Added %d new row(s). pac_review now has %d rows total.\n",
              nrow(new_rows), nrow(pac_review)))
}

cat(sprintf("\nFinal pac_review: %d rows, %d with no pac_id (name-matched)\n",
            nrow(pac_review),
            sum(is.na(pac_review$pac_id) | pac_review$pac_id == "")))
```

```{r}
pac_review %>% filter(pac_id == "1370549")
```

```{r}
pac_review %>% filter(Contributor.Name == "Service Employees International Union Committee of Interns and Resident Physicians")
```

also look for industry_override is major/mega donor...

### post-merge: fix federal flags and propagate overrides by fec_id

```{r}
# 1. propagate fec_id within pac_id groups
#    same pac_id can span multiple rows where only one row got a coded fec_id from review
fec_by_pac_id <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "", !is.na(fec_id), fec_id != "") %>%
  group_by(pac_id) %>%
  slice(1) %>%
  ungroup() %>%
  select(pac_id, fec_id_fill = fec_id)

pac_review <- pac_review %>%
  left_join(fec_by_pac_id, by = "pac_id") %>%
  mutate(fec_id = coalesce(fec_id, fec_id_fill)) %>%
  select(-fec_id_fill)

# 2. any row with a non-NA fec_id is a federal PAC
pac_review <- pac_review %>%
  mutate(is_federal_pac = if_else(
    !is.na(fec_id) & fec_id != "",
    "TRUE", is_federal_pac
  ))

# 2b. propagate is_federal_pac = "TRUE" across all rows in the same pac_id group
fed_pac_ids <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "", coalesce(is_federal_pac, "") == "TRUE") %>%
  distinct(pac_id)

pac_review <- pac_review %>%
  mutate(is_federal_pac = if_else(
    !is.na(pac_id) & pac_id != "" & pac_id %in% fed_pac_ids$pac_id,
    "TRUE", is_federal_pac
  ))

# 3. propagate industry override within fec_id groups
#    if multiple pac_ids share a fec_id, they should have consistent overrides.
#    where there is exactly one distinct non-PENDING override -> fill into rows that are missing/PENDING.
#    where overrides genuinely conflict -> warn and print for manual review.
fec_canonical <- pac_review %>%
  filter(!is.na(fec_id), fec_id != "",
         !is.na(industry_override_under_5k),
         industry_override_under_5k != "",
         industry_override_under_5k != "PENDING") %>%
  group_by(fec_id) %>%
  summarize(
    n_distinct = n_distinct(industry_override_under_5k),
    canonical  = dplyr::first(industry_override_under_5k),
    .groups = "drop"
  )

fec_id_conflicts <- fec_canonical %>% filter(n_distinct > 1)
if (nrow(fec_id_conflicts) > 0) {
  warning(sprintf(
    "%d fec_id group(s) have conflicting industry overrides across pac_ids — manual review needed.",
    nrow(fec_id_conflicts)
  ))
  print(pac_review %>%
    filter(fec_id %in% fec_id_conflicts$fec_id) %>%
    select(fec_id, pac_id, Contributor.Name, industry_override_under_5k) %>%
    arrange(fec_id))
}

fec_propagate <- fec_canonical %>%
  filter(n_distinct == 1) %>%
  select(fec_id, fec_canonical_override = canonical)

pac_review <- pac_review %>%
  left_join(fec_propagate, by = "fec_id") %>%
  mutate(industry_override_under_5k = if_else(
    !is.na(fec_id) & fec_id != "" & !is.na(fec_canonical_override) &
    (is.na(industry_override_under_5k) | industry_override_under_5k == "" |
     industry_override_under_5k == "PENDING"),
    fec_canonical_override,
    industry_override_under_5k
  )) %>%
  select(-fec_canonical_override)

cat(sprintf(
  "Rows with fec_id but is_federal_pac != 'TRUE' remaining: %d\n",
  sum(!is.na(pac_review$fec_id) & pac_review$fec_id != "" &
      coalesce(pac_review$is_federal_pac, "") != "TRUE", na.rm = TRUE)
))
cat(sprintf(
  "fec_id override propagation: %d unambiguous fec_id group(s) propagated, %d conflict(s) flagged\n",
  nrow(fec_propagate), nrow(fec_id_conflicts)
))
```

```{r}
contributions %>% filter(Recipient.Name == "ALLEN, BEN", Office == "Insurance Commissioner") %>% select(Recipient.Committee,Recipient.Committee.ID)
```


### deduplicate pac_review to one row per pac_id

```{r}
# check for conflicting industry overrides across duplicate rows before collapsing
pac_id_override_conflicts <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  group_by(pac_id) %>%
  summarise(
    n_valid_overrides = n_distinct(industry_override_under_5k[
      !is.na(industry_override_under_5k) &
      industry_override_under_5k != "" &
      industry_override_under_5k != "PENDING"
    ]),
    conflicting_values = paste(sort(unique(industry_override_under_5k[
      !is.na(industry_override_under_5k) &
      industry_override_under_5k != "" &
      industry_override_under_5k != "PENDING"
    ])), collapse = " | "),
    .groups = "drop"
  ) %>%
  filter(n_valid_overrides > 1)

if (nrow(pac_id_override_conflicts) > 0) {
  warning(sprintf(
    "%d pac_id(s) have conflicting industry overrides across duplicate rows — manual resolution needed.",
    nrow(pac_id_override_conflicts)
  ))
  print(pac_review %>%
    filter(pac_id %in% pac_id_override_conflicts$pac_id) %>%
    select(pac_id, Contributor.Name, industry_override_under_5k, is_federal_pac, fec_id) %>%
    arrange(pac_id))
} else {
  cat("No industry override conflicts across duplicate pac_id rows.\n")
}

# collapse to one row per pac_id, combining best values across duplicates
pac_review_noid <- pac_review %>% filter(is.na(pac_id) | pac_id == "")

pac_review <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  group_by(pac_id) %>%
  summarise(
    across(everything(), first),
    # "TRUE" wins if any row in the group has it
    is_federal_pac = if_else(
      any(coalesce(is_federal_pac, "") == "TRUE"), "TRUE", first(is_federal_pac)
    ),
    # take first non-NA, non-empty, non-PENDING override; fall back to first value
    industry_override_under_5k = {
      valid <- industry_override_under_5k[
        !is.na(industry_override_under_5k) &
        industry_override_under_5k != "" &
        industry_override_under_5k != "PENDING"
      ]
      if (length(valid) > 0) valid[1] else first(industry_override_under_5k)
    },
    .groups = "drop"
  ) %>%
  bind_rows(pac_review_noid)

# merge no-id rows into id-matched rows where name matches (case-insensitive)
id_names_upper <- pac_review %>%
  filter(!is.na(pac_id), pac_id != "") %>%
  mutate(name_key = toupper(trimws(Contributor.Name))) %>%
  select(matched_pac_id = pac_id, name_key)

noid_candidates <- pac_review %>%
  filter(is.na(pac_id) | pac_id == "") %>%
  mutate(name_key = toupper(trimws(Contributor.Name))) %>%
  inner_join(id_names_upper, by = "name_key") %>%
  select(-pac_id) %>%
  rename(pac_id = matched_pac_id)

if (nrow(noid_candidates) > 0) {
  cat(sprintf("%d no-id row(s) matched to an id-matched row by name — merging and removing duplicates:\n",
              nrow(noid_candidates)))
  print(noid_candidates %>%
    select(name_key, pac_id, industry_override_under_5k, is_federal_pac, fec_id))

  # fill missing override on id row from no-id row where id row has none
  noid_fill <- noid_candidates %>%
    filter(!is.na(industry_override_under_5k),
           industry_override_under_5k != "",
           industry_override_under_5k != "PENDING") %>%
    distinct(pac_id, .keep_all = TRUE) %>%
    select(pac_id, noid_override = industry_override_under_5k)

  pac_review <- pac_review %>%
    left_join(noid_fill, by = "pac_id") %>%
    mutate(industry_override_under_5k = if_else(
      !is.na(pac_id) & pac_id != "" & !is.na(noid_override) &
      (is.na(industry_override_under_5k) | industry_override_under_5k == "" |
       industry_override_under_5k == "PENDING"),
      noid_override, industry_override_under_5k
    )) %>%
    select(-noid_override)

  # drop the no-id rows that were merged into id rows
  merged_name_keys <- noid_candidates %>% pull(name_key) %>% unique()
  pac_review <- pac_review %>%
    filter(
      !is.na(pac_id) & pac_id != "" |
      !toupper(trimws(Contributor.Name)) %in% merged_name_keys
    )
} else {
  cat("No no-id rows matched to id-matched rows by name.\n")
}

cat(sprintf(
  "pac_review after dedup: %d rows (%d with pac_id, %d without)\n",
  nrow(pac_review),
  sum(!is.na(pac_review$pac_id) & pac_review$pac_id != ""),
  sum(is.na(pac_review$pac_id) | pac_review$pac_id == "")
))
```

### data checks (can skip to next section)

```{r}
# major donors
bind_rows(all_major_donors)

# inspect conflicts
bind_rows(all_id_conflicts)

# inspect name conflicts
bind_rows(all_name_conflicts)

# inspect newly resolved pacs
bind_rows(all_pending_resolved)
```

```{r}
# make sure fed flag persisted
fed_ids <- supplemental_reviews %>% filter(is_federal_pac == "TRUE", !is.na(pac_id)) %>% pull(pac_id)
pac_review %>% filter(pac_id %in% fed_ids) %>% distinct(is_federal_pac)

#pac_review %>% filter(!is.na(fec_id) & coalesce(is_federal_pac, "") != "TRUE") #%>% distinct(is_federal_pac)
#pac_review %>% filter(!is.na(pac_id)) %>% group_by(pac_id) %>% filter(n_distinct(industry_override_under_5k) > 1)
pac_review %>% filter(!is.na(fec_id)) %>% group_by(fec_id) %>% filter(n_distinct(industry_override_under_5k) > 1)
#pac_review %>% filter(!is.na(pac_id)) %>% group_by(pac_id) %>% filter(n_distinct(fec_id) > 1 | n_distinct(is_federal_pac) > 1) %>% arrange(pac_id)
```

# Flag major donors

```{r}
pac_review <- pac_review %>%
  mutate(is_major_donor =
    trimws(coalesce(industry_override_under_5k, "")) == "Major Donors/Megadonors" |
    grepl("major donor", coalesce(notes, ""), ignore.case = TRUE)
  )

cat(sprintf("pac_review: %d row(s), %d flagged as major donor\n",
            nrow(pac_review), sum(pac_review$is_major_donor, na.rm = TRUE)))
```

# Write output

```{r}
write_csv(pac_review, "04_inputs/pac_manual_review_reconciled.csv")
cat(sprintf("pac_manual_review_reconciled.csv written: %d rows\n", nrow(pac_review)))
```

